# Exercises

## Unix

:::{.notice}

Following is based on the “SciServer Essentials 2.0” image described in Ch 2. Setup Compute

:::

### Run a command

- Start a terminal

```{r, echo = FALSE}

ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_0")

```

Run the `ls -l` command to list files (the `-l` is a command line argument that instructs the `ls` program to modify its operation so that longer details are provided about each file)

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_8")
```


### Wrap a command in a Bash

- Create Text File

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_29")
```

- Write your first Bash script

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_40")
```


- Run `00-hello.sh`

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_45")
```



## Python

### Learning objectives

- Understand how to run Python scripts from the command line

- Use a terminal to navigate directories and execute Python code

### Hello world example

1. Open the terminal.

2. Navigate to the directory where you want to create your Python script.

3. Create a new file named `hello.py` using a text editor of your choice (e.g., `nano`, `vim`, or `touch`).

4. Open `hello.py` in the text editor and write the following code:

   ```python
   print("Hello, world!")
   ```

5. Save the file and exit the text editor.

6. In the terminal, run the script by typing:

   ```bash
   python hello.py
   ```

7. You should see the output:

   ```
   Hello, world!
   ```



## Command line arguments

### import sys

### sys.argv

### lists

### types

## Parsing file line by line

### File streams

### for loops

## head.py

## grep.py

## cut.py