```{r echo = FALSE}
knitr::opts_chunk$set(out.width = "100%")
```



# Running Programs

:::{.notice}

Following is based on the “SciServer Essentials 2.0” image described in Ch 2. Setup Compute

:::

## Learning objectives

- Run a UNIX command

- Wrap a command in a Bash script with `#!`

- Make a Bash script executable


## Run a command

- Start a terminal

```{r, echo = FALSE}

ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_0")

```

Run the `ls -l` command to list files (the `-l` is a command line argument that instructs the `ls` program to modify its operation so that longer details are provided about each file)

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_8")
```


## Wrap a command in a Bash

- Create Text File

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_29")
```

- Write your first Bash script

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_40")
```


- Run `00-hello.sh`

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_45")
```



# Hello, World!

## Learning objectives

- Understand how to run Python scripts from the command line

## Hello world example

- Follow the steps from the ***SciServer*** lesson
    
    - Login to SciServer and start a new jupyter notebook

    - Start a new terminal session
    

- Create a new file named `01-helloworld.py` using the text editor and write the following code:

    ```python
    #!/usr/bin/env python3
    print("Hello, world!")
    ```

- Save the file and make it executable:

    ```bash
    chmod +x 01-helloworld.py
    ```

- In the terminal, run the script by typing:

    ```bash
    ./01-helloworld.py
    ```

- You should see the output:

   ```
   Hello, world!
   ```

Congratulations! You have just:

- Created a Python script that prints "Hello, world!" to the terminal

- Made the script executable

- Ran the script from the command line

# Command line arguments

## Printing All Command Line Arguments

Let's make our Python scripts interactive by accepting input from the command line. Create a new file called `02-arguments.py` and type:

```python
#!/usr/bin/env python3

import sys

print(sys.argv)
```

Save the file and make it executable:

```bash
chmod +x 02-arguments.py
```

Now run it directly with some arguments:

```bash
./02-arguments.py hello world 123
```

You'll see output like `['./02-arguments.py', 'hello', 'world', '123']`

The `sys` module provide access to the system-specific parameter. The variable `sys.argv` contains all command line arguments passed to your script, including the script name itself as the first element.


## Understanding Lists

Lists in Python are ordered collections of items enclosed in square brackets, like `[1, 2, 3]` or `["apple", "banana"]`. Lists can contain different types of data and are accessed by the position (index), starting from 0. For example, `my_list[0]` gets the first item, `my_list[1]` gets the second item, and so in.


## Accessing Specific Arguments

Lets modify our script to print just the second command line argument:

```python
#!/usr/bin/env python3

import sys

print("Script name:", sys.argv[0])
print("First argument:", sys.argv[1])
```

Run it with: `./02-arguments.py hello`

This prints:

```bash
Script name: ./02-arguments.py
First argument: hello
```

Notice how `sys.argv[0]` is always the script name, so the first actual argument is at index 1.

***Warning***: If you don't provide enough arguments, Python will crash with an "IndexError". We'll learn to handle this with `if` statements later.

## Arguments Are Strings

Command line arguments are always strings, even if they look like numbers. Update `02-arguments.py`:

```python
#!/usr/bin/env python3

import sys

# This won't work as expected
result = sys.argv[1] + sys.argv[2]
print("Without conversion:", result)

# Convert strings to integers first
num1 = int(sys.argv[1])
num2 = int(sys.argv[2])
print("With conversion:", num1 + num2)
```

Run it with: `./02-arguments.py 5 3`

Output:

```bash
Without conversion: 53
With conversion: 8
```

Without conversion, Python concatenates the strings "5" and "3" into "53". The `int()` function converts string representations of numbers into actual integers that can be used in mathematical operations.

Congratulations! You have just:

- Created a Python script that accepts command line arguments

- Accessed specific arguments using `sys.argv`




# Parsing files line by line

## Create a Test File

Let's create a simple text file called `sample.txt` using the Jupyter text editor, or create it directly from the command line.

```bash
echo -e "apple\nbanana\ncherry\ndate" > sample1.txt
```

This creates a file with five fruit names, each on its own line.

## File Streams in Python

A file stream is like a pipeline that lets you read data from a file one piece at a time. The omst common way to open a file is using the `open()` function.

- Create a new Python script called `03-files.py`

```python
#!/usr/bin/env python3

import sys

my_file = open( sys.argv[1] )
print(my_file)

```

- Save it and make it executable

```bash
chmod +x 03-files.py
``` 

- Run the script with the file name as an argument.

```bash
./03-files.py sample.txt
``` 

This will print something like

```bash
<_io.TextIOWrapper name='sample1.txt' mode='r' encoding='UTF-8'>
```

As you can see, the `print()` function can't print the file content directly. This output just indicates the file `sample1.txt` is opened in read mode (`'r'`) with UTF-8 encoding.

## `for` loops

To read the file and print each line, we can use a `for` loop. 

- Update `03-files.py`:

```python
#!/usr/bin/env python3

import sys

my_file = open(sys.argv[1])
for line in my_file:
    print(line.strip())
```

- Save it and run the script again

```bash
./03-files.py sample1.txt
```

This will print each line of the file without extra spaces or newlines

```bash
apple
banana
cherry
date
```

We can also use for loops to read multiple files. For example, to read files `sample1.txt` and `sample2.txt`.

- create a new file called `sample2.txt` with the following content

```bash
echo -e "fig\nkiwi\nlemon" > sample2.txt
```

- Update `03-files.py` to read both files

```python
#!/usr/bin/env python3

import sys

# for iterates through files
for my_line in my_file:
    # objects have methods
    my_line = my_line.rstrip("\n")
    print( my_line )

my_file.close()
```
- Save it and run the script with both files

```bash
./03-files.py sample1.txt sample2.txt
```

This will print all lines from both files:

```bash
apple
banana
cherry
date
fig
kiwi
lemon
``` 

Congratulations! You have just:

- Created a Python script that reads a file 

- Used a `for` loop to iterate through each line in the file

- Printed each line without extra spaces or newlines


# head.py

# grep.py

# cut.py