```{r echo = FALSE}
knitr::opts_chunk$set(out.width = "100%")
```



# Running Programs

:::{.notice}

Following is based on the “SciServer Essentials 2.0” image described in Ch 2. Setup Compute

:::

## Learning objectives

- Run a UNIX command

- Wrap a command in a Bash script

- Make a Bash script executable


## Run a command

- Start a terminal

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_0")
```

- Run the `ls -l` command to list files (the `-l` is a command line argument that instructs the `ls` program to modify its operation so that longer details are provided about each file)

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_8")
```

- Run `echo "Hello, World!"` to print text to the terminal

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit?slide=id.g36ef0094ecf_1_0#slide=id.g36ef0094ecf_1_0")
```

## Wrap a command in a Bash

- Create Text File

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_29")
```

- Write your first Bash script

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_40")
```


- Run `00-hello.sh`

```{r, echo = FALSE}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g36ef0094ecf_0_45")
```


Congratulations! You have just:

- Created a Bash script that prints "Hello, World!" to the terminal


# Hello, World!

## Learning objectives

- Understand how to run Python scripts from the command line

## Hello, World! example

- Follow the steps from the ***SciServer*** lesson
    
    - Login to SciServer and start a new jupyter notebook

    - Start a new terminal session
    

- Create a new file named `01-helloworld.py` using the text editor and write the following code:

    ```python
    #!/usr/bin/env python3
    print("Hello, World!")
    ```

    - The beginning of the first line, `#!`,  it is called a "hashbang" or "shebang". That indicates which interpreted should process (Python in this case) the file.

    - We use `/usr/bin/env` to find the Python interpreter in the user's environment, which is more flexible than specifying a path like `/usr/bin/python3`. This way, it works regardless of where Python is installed on the system.

    - `python3` is the interpreter that will run the script. It ensures that the script is executed with Python 3, which is important because Python 2 and Python 3 have different syntax and features.

    - Lastly `print("Hello, World!")` is the actual Python code that prints "Hello, World!" to the terminal.

    

- Save the file and make it executable:

    ```bash
    chmod +x 01-helloworld.py
    ```

- In the terminal, run the script by typing:

    ```bash
    ./01-helloworld.py
    ```

- You should see the output:

   ```
   Hello, World!
   ```

Congratulations! You have just:

- Created a Python script that prints "Hello, World!" to the terminal

- Made the script executable

- Ran the script from the command line

# Command line arguments

## Learning objectives

- Understand how to accept command line arguments in Python scripts

- Learn how to access command line arguments using the `sys` module

- Understand how to handle command line arguments

## Printing All Command Line Arguments

- Let's make our Python scripts interactive by accepting input from the command line. Create a new file called `02-arguments.py` and type:

    ```python
    #!/usr/bin/env python3

    import sys

    print(sys.argv)
    ```

- Save the file and make it executable:

    ```bash
    chmod +x 02-arguments.py
    ```

- Now run it directly with some arguments:

    ```bash
    ./02-arguments.py hello world 123
    ```

    You'll see output like `['./02-arguments.py', 'hello', 'world', '123']`

The `sys` module provide access to the system-specific parameter. The variable `sys.argv` contains all command line arguments passed to your script, including the script name itself as the first element.


## Understanding Lists

Lists in Python are ordered collections of items enclosed in square brackets, like `[1, 2, 3]` or `["apple", "banana"]`. Lists can contain different types of data and are accessed by the position (index), starting from 0. For example, `my_list[0]` gets the first item, `my_list[1]` gets the second item, and so in.


## Accessing Specific Arguments

- Lets modify our script to print just the second command line argument:

    ```python
    #!/usr/bin/env python3

    import sys

    print("Script name:", sys.argv[0])
    print("First argument:", sys.argv[1])
    ```

- Run it with: `./02-arguments.py hello`

    This prints:

    ```bash
    Script name: ./02-arguments.py
    First argument: hello
    ```

    Notice how `sys.argv[0]` is always the script name, so the first actual argument is at index 1

    ***Warning***: If you don't provide enough arguments, Python will crash with an "IndexError". We'll learn to handle this with `if` statements later

## Arguments Are Strings

Command line arguments are always strings, even if they look like numbers. 

- Update `02-arguments.py`

    ```python
    #!/usr/bin/env python3

    import sys

    # This won't work as expected
    result = sys.argv[1] + sys.argv[2]
    print("Without conversion:", result)

    # Convert strings to integers first
    num1 = int(sys.argv[1])
    num2 = int(sys.argv[2])
    print("With conversion:", num1 + num2)
    ```

- Run it with: `./02-arguments.py 5 3`

    Output:

    ```bash
    Without conversion: 53
    With conversion: 8
    ```

Without conversion, Python concatenates the strings "5" and "3" into "53". The `int()` function converts string representations of numbers into actual integers that can be used in mathematical operations.

Congratulations! You have just:

- Created a Python script that accepts command line arguments

- Accessed specific arguments using `sys.argv`


# Parsing files line by line

## Learning objectives

- Understand how to read files in Python

- Learn how to iterate through lines in a file using a `for` loop

## Create a Test File

- Let's create a simple text file called `sample.txt` using the Jupyter text editor

    The file should contain the following lines:

    ```
    apple
    banana
    cherry
    date

    ```

## File Streams in Python

A file stream is like a pipeline that lets you read data from a file one piece at a time. The most common way to open a file is using the `open()` function.

- Create a new Python script called `03-files.py`

    ```python
    #!/usr/bin/env python3

    import sys

    my_file = open( sys.argv[1] )
    print(my_file)

    ```

- Save it and make it executable

    ```bash
    chmod +x 03-files.py
    ``` 

- Run the script with the file name as an argument

    ```bash
    ./03-files.py sample.txt
    ``` 

    This will print something like

    ```bash
    <_io.TextIOWrapper name='sample1.txt' mode='r' encoding='UTF-8'>
    ```

As you can see, the `print()` function can't print the file content directly. This output just indicates the file `sample1.txt` is opened in read mode (`'r'`) with UTF-8 encoding.

## `for` loops

To read the file and print each line, we can use a `for` loop. 

- Update `03-files.py`:

    ```python
    #!/usr/bin/env python3

    import sys

    my_file = open(sys.argv[1])

    # for iterates through files
    for my_line in my_file:
        # objects have methods
        my_line = my_line.rstrip("\n")
        print( my_line )

    my_file.close()
    ```
    - We use `my_file.close()` to close the file after we're done reading it. This is important to free up system resources.

- Save it and run the script again

    ```bash
    ./03-files.py sample1.txt
    ```

    This will print each line of the file without extra spaces or newlines

    ```bash
    apple
    banana
    cherry
    date
    ```

Congratulations! You have just:

- Created a Python script that reads a file 

- Used a `for` loop to iterate through each line in the file

- Printed each line without extra spaces or newlines


# Replicating the `head` command in Python

## Learning objectives

- How to replicate the `head` bash command in Python

- Understand how to read a file and print the first few lines

- Learn how to use a `for` loop to limit the number of lines printed

## Create a new file

- Let's create a simple text file called `sample2.txt` using the Jupyter text editor

    The file should contain the following lines:

    ```
    apple
    banana
    cherry
    date
    elderberry
    fig
    grape
    peach
    kiwi
    lemon
    ```

## Practice the `head` command

The `head` command in Bash prints the first few lines of a file. By default, it shows the first 10 lines, but you can specify a different number with the `-n` option.

- In the terminal, run the command:

    ```bash
    head sample2.txt
    ```
    
    - This will print all the 10 lines of `sample2.txt`

        ```
        apple
        banana
        cherry
        date
        elderberry
        fig
        grape
        peach
        kiwi
        lemon
        ```
- Now try running it with the `-n` option to print only the first 5 lines

    ```bash
    head -n 5 sample2.txt
    ```
    - You should see the following output

        ```
        apple
        banana
        cherry
        date
        elderberry
        ```

## Replicate the `head` command in Python

Now let's create a Python script that replicates the `head` command.

- Create a new Python script called `04-head.py`

    ```python
    #!/usr/bin/env python3

    import sys

    my_file = open( sys.argv[1] )

    max_lines = 10
    if len(sys.argv) > 2:
        max_lines = int(sys.argv[2])

    # be careful of off-by-1 errors
    i = 0
    for my_line in my_file:
        if i >= max_lines:
            break
        my_line = my_line.rstrip("\n")
        print( my_line )
        i = i + 1

    my_file.close()
    ```

    - The script takes two command line arguments: 
        - the file name `sys.argv[1]`
        
        - The number of lines to print `sys.argv[2]`
        
        `If` the second argument is not provided, it defaults to 10.
        
        - The `If` statement is used to check if the number of lines exceeds the maximum specified.

- Save the file and make it executable

    ```bash
    chmod +x 04-head.py
    ```

- Run the script with the file name and number of lines as arguments

    ```bash
    ./04-head.py sample2.txt 5
    ```

    This will print the first 5 lines of `sample2.txt`

    ```
    apple
    banana
    cherry
    date
    elderberry
    ```

Congratulations! You have just:

- Created a Python script that replicates the `head` command

- Used command line arguments to specify the file and number of lines to print

# grep.py

# cut.py