---
title: "Zero-to-Python"
date: "`r format(Sys.time(), '%B, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: "Description about Course/Book."
favicon: assets/dasl_favicon.ico
output:
    bookdown::word_document2:
      toc: true
---

```{r echo = FALSE}
knitr::opts_chunk$set(out.width = "100%")
```


# About this Course {-}

This course serves as first introduction to Python by introducing how to write scripts using a text editor and runnning the, at the UNIX command line.
By the end of the exercises you should have the tools you need to follow and full fledged Python tutorial.

```{r, echo=FALSE, fig.alt=''}
ottrpal::include_slide("https://docs.google.com/presentation/d/1Covg_bpPYGKnTVbOjX_FkGgWHBG6VwdEySKbpZNwqKo/edit#slide=id.g3710e1bbefd_0_16")
```

### Learning Goals {-}

- Practice a writting script, running it, updating script, running it, creating new script.

- Parse text files line by line.

- Appreciate how Unix commands are just files/programs/scripts.